An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets

### 2021-06-09 CMPSpaper_writing/CMPSpaper_writing.Rmd
added codes and data for reproducibility
removed appendix; switched the usage of reference and comparison; added link to the github repo

The white horizontal line in \autoref{fig:process}(a) indicates which cross-section was identified by the algorithm to represent the LEA; \autoref{fig:process}(b) shows the corresponding cross-sectional view. 
Groove engraved areas are removed from the analysis as indicated by the vertical blue lines in \autoref{fig:process}(c) and \autoref{fig:process}(d)). A non-parametric LOESS smooth  [@loess] is fitted to capture the bullet curvature (\autoref{fig:process}(e)) and, finally, the **bullet signature** (\autoref{fig:process}(f)) is obtained as residuals of the cross-section and the fitted smooth. 
Note, that in @cmps bullet signatures are  referred to as bullet profiles. However, to avoid confusion, we distinguish the notion of bullet signatures from bullet profiles. Bullet profiles are shown in panels (b), (c), and (d) of \autoref{fig:process}, while \autoref{fig:process}(f) shows the corresponding bullet signature.
Identifiying the groove engraved areas correctly is fundamental for a correct down-stream analysis of the signatures. To allow for a human-in-the middle inspection and intervention we have provided an interactive web-application, implemented as `R` Shiny App [@shiny],  `bulletinspectR` to identify and correct those errors. 
\hh{An example of the extraction process with corresponding code and parameter settings can be found at XXX.}
Note that the process of extracting signatures might be different from the one used in @cmps because no code or parameter settings are made available publicly.

@@

### Conceptual idea of CMPS
Most algorithms for comparing striation marks are based on the digitized signatures and produce a similarity score [@song2005, @ChumbleyL_Scott2010VoTM, @aoas, @pmid30444940 ].
 The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm. The algorithm's main idea  is to take a set of consecutive and non-overlapping basis segments from  the reference   and  for each segment find the "best" registration position on the comparison  (the other bullet signature) with respect to their cross-correlation values. From a comparison of these registration positions, a **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score. 
  High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair. Low scores between pairs of signatures are attributed to different source pairs.
\wj{However, a specific threshold of the CMPS score to distinguish between same-source and different-source comparison is not mentioned in @cmps and depends on the underlying structure of the data and the choice of parameters. Further research is needed to understand how to determine the threshold. XXX}
  <!-- \hh{XXX What threshold did @cmps identify as the boundary between different source and same source? And what was the error rate in their example?} -->
  <!-- \wj{they didn't talk about a specific threshold or error rates} \hh{XXX Is there a way to infer the error based on the example? ... 'while the authors did not mention a specific threshold of the CMPS to distinguish between same-source and different-source comparison, a CMPS value of Y results in an error rate of Z in the example.'} -->
 The CMPS algorithm can assist firearm examiners with drawing a conclusion about the source of a comparison pair. Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the CMPS algorithm. 
Thus, in this paper we present an open-source implementation of the CMPS algorithm \hh{in the R package CMPS available from XXX github / CRAN}. 
With our implementation, one can compute the CMPS score of a comparison using the following code:

@@

1. **cut the reference signature into consecutive, non-overlapping and equal-length basis segments:**
The command `get_segs(x, len=50)` implements this step:  it takes  bullet signature `x` in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`, which are referred to as "basis segments".

2. **identify candidate positions:** For each basis segment a set of candidate registration  positions on the comparison signature is identified based on the segment's similarity to the comparison. In a first step, the cross-correlation function of the segment to the comparison is calculated, then a number of positions with high correlation values are identified as candidate positions. In case multiple segment lengths are considered, the length of each basis segment is doubled and these two steps are repeated. Only when candidate positions coincide (or are similar enough), they are considered further. \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.
    + **Calculate the cross-correlation curve:**  Calculate the cross-correlation curve between a basis segment `x` and the comparison signature `y` using the function `get_ccf4(x, y, ...)` as shown in \autoref{fig:segplot2}. The position indicates the lag by which a basis segment is moved with respect to its original placement. A position is considered "good" if it results in a peak in the cross-correlation between the basis segment and the comparison.

@@

        -   `get_ccr_peaks(comp, segments, seg_scale, nseg = 1, npeaks = 5)` computes the cross-correlation curve between a basis or scaled segment and the comparison signature and finds peaks in the cross-correlation curve. The number of peaks detected is equal to `npeaks`, which is an integer. `segments`, `seg_scale`, and `nseg` determines the segment in the cross-correlation computation, and `comp` gives the comparison signature. If the multi-segment lengths strategy is used, then `get_ccr_peaks(...)` is called in a `lapply()` for each level of the basis segment. The resulting list is called `ccr.list`.

@@

3. **determine the congruent registration position:** A candidate position "receives" `votes` from basis segments that identify it or a close position within a tolerance zone of `Tx` as a candidate position in step 2. Votes for all candidate positions are tallied, and the position with the highest number of votes gets chosen as the *congruent registration position*,  indicating that most of the basis segments find their highly similar counterpart in the comparison signature in terms of correlation at this registration position. In case of ties the middle position is taken as the congruent registration position. Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score of the comparison. `get_CMPS(input.ccp, Tx = 25)` is the function that tallies the votes and determines the congruent registration position and congruent matching profile segments (cmps).

@@

* `x` and `y` are  two  signatures:  `x` serves as the reference signature (which will be divided into basis segments) and `y` is the comparison signature;

* `include` determines the \hh{format of the function result}. Besides the `CMPS.score`, other aspects of the comparison help in understanding how the `CMPS.score` is computed. By default `include` is set to `NULL` and  only  the `CMPS.score` is returned; further results are included when `include` is (an abbreviation of) one of or a vector of the following strings: `"nseg", "congruent.pos", "congruent.seg.idx", "segments", "parameters"`, and `"full_result"`. If `include` is specified as `"full_result"` (or its abbreviation), the ouput includes everything listed below.
  - `nseg`: the number of basis segments  from the reference signature; this is also the highest possible CMPS score of the comparison;
  - `congruent.pos`: the congruent registration position;
  - `congruent.seg.idx`: the indices of all congruent matching profile segments;
  - `ccp.list`: a list showing identified candidate positions of all basis segments; 
  - `pos.df`: a data frame containing all candidate positions and their respective number of votes;
  - `segments`: a list containing all basis segments;
  - `parameters`: a list containing all input arguments of `extract_feature_cmps`;

@@

The `CMPS` package contains a simple example to illustrate the basic usage of the package. The data in this example are twelve bullet signatures obtained from two bullets in Hamby set 252 [@hamby].  The procedure for generating signatures from high-resolution 3D topographic scans of bullet lands used here follows the methodology described in @aoas (as discussed above). 
The two bullets under consideration are known to have been fired from the same gun barrel, so for the 36 pairwise land-by-land comparisons, six comparison are from same-source pairs (known matches) while thirty are from different-source pairs (known non-matches). To access the example data, we use

@@

The signatures of Land 4 of Bullet 1 and Land 5 of Bullet 2 are stored in objects `sigs2` and `sigs1`, respectively. This comparison is one of a pair of signatures that are known to be a match -- a KM (known match) comparison. We compute the CMPS score using two versions of the CMPS algorithm:

@@

We also implemented graphing tools for visualizing results of the CMPS algorithm. The goal is to provide users with tools to inspect each of the basis segments and to help them have a better understanding of how the algorithm works. \autoref{fig:sigplots} shows the plots generated by the first graphing function, `cmps_signature_plot()`, and continues with the example above. `cmps_signature_plot()` takes the output of `extract_feature_cmps(..., include = "full_result")` and returns a list of 5 elements. It creates an overall impression of how the reference signature aligns with the comparison signature at the congruent registration position.

@@

*   The second plot is called `signature_shift_plot`, shown in \autoref{fig:sigplot2}. This visual presents both the reference signature and the comparison signature. The comparison signature is aligned with the reference signature based on the congruent registration position. Congruent matching profile segments are highlighted by  solid red lines.

\begin{figure}[hbt]
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot}The black line shows the comparison signature; each red line segment shows one congruent matching profile segment.}
```{r sigplot, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
sig.plot <- cmps_signature_plot(
  cmps_with_multi_scale
)
sig.plot$segment_shift_plot
```
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot2}The black line shows the comparison signature; the red line shows the reference signature. Solid part shows the congruent matching profile segments, and the dashed part shows segments that do not agree with the congruent registration position.}
```{r sigplot2, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
sig.plot$signature_shift_plot
```
\end{subfigure}
\caption{\label{fig:sigplots} Two visual outputs of cmps\_signature\_plot()}
\end{figure}

@@

*   \autoref{fig:segplot} is the `segment_plot` for basis segment 6 at level one (in its original length). We used  `npeaks.set = c(5, 3, 1)` in `extract_feature_cmps()` when calculating the CMPS score. Therefore the top five peaks are identified in the cross-correlation curve at level one. Segment 6 is plotted at the positions where these 5 peaks are identified with dashed lines in the `segment_plot`. The solid thick black line shows the segment at its original position (which in this example is very close to the actual registration position). 
*   \autoref{fig:segplot2} is the `scale_ccf_plot` of basis segment 6 at level one. It shows the cross-correlation curve computed by the comparison signature and the level-one basis segment 6. The five highest peaks are marked by dots on the curve.

\begin{figure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:segplot}segment\_plot for segment 6 at level one. The original position of segment 6 is indicated by the solid black line. Positions where the segment achieves the 5 highest cross-correlations are indicated by the dashed line segments.}
```{r segplot, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
seg.plot <- cmps_segment_plot(
  cmps_with_multi_scale, 
  seg.idx = 6
)
seg.plot[[1]]$segment_plot
```
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:segplot2}scale\_ccf\_plot shows the cross-correlation curve between the comparison signature and segment 6 at level one. The five highest peaks are marked by dots. The vertical red dashed line indicates the congruent registration position; the green dashed line shows a peak position in the highest segment level; the blue dashed lines show the tolerance zone around the green dashed line. We can see that none of the five highest peaks at level one falls within the tolerance zone, indicating that there is no consistent correlation peak or a candidate position identified by basis segment 6 under the multi-segment lengths strategy. Thus, the basis segment 6 doesn't vote for the congruent registration position and is not a cmps.}
```{r segplot2, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
seg.plot[[1]]$scale_ccf_plot
```
\end{subfigure}
\caption{\label{fig:segplots} Two plots used to investigate basis segment 6 at level one}
\end{figure}

@@

ggarrange(
  plotlist = 
    unlist(seg.plot, 
           recursive = FALSE),
  ncol = 2, 
  nrow = 3)
```
\caption{\label{fig:seg_all} Put segment\_plot and scale\_ccf\_plot of all three levels together. We are identifying five highest peaks at level one, three peaks at level two, and one peak at level three since `npeaks.set = c(5, 3, 1)`. The highest peak position at level three is marked by the green dashed line across all segment levels. However, the highest peak on level three does not coincide with any of the top five highest peaks at level one. This indicates that there is no consistent correlation peak or a candidate position for basis segment 6 under the multi-segment lengths strategy.}
\end{figure}

@@

In general, we assume each bullet has $n$ land engravings (in our case $n=6$). Let $c_{ij}$ denote the CMPS score of comparison between bullet 1 land $i$ and bullet 2 land $j$, for $i,j = 1, \dots, n$. Let $\mathcal{P}_k$ denote bullet land pairs in phase $k$ for $k = 0, \dots, n-1$, and $\mathcal{P}_k = \{ (i,j): i = 1, \dots, n ; j = (i + k) - n \cdot \mathbbm{1}_{i + k > n} \}$, where $\mathbbm{1}_A$ denotes an indicator function. For example, $\mathcal{P}_1 = \{ (1,2), (2,3), (3,4), (4,5), (5,6), (6,1) \}$ when $n = 6$.

@@

\wj{The codes generating those results can be found in the appendix.XXX}

@@

<!-- <div id="refs"></div> -->

\newpage

# Appendix

```{r generate252, echo=TRUE, eval=FALSE}
library(tidyverse)
library(bulletxtrctr)
library(x3ptools)
library(CMPS)
library(ggpubr)
library(parallel)

###########
# from the x3p data, re-run CMPS results for hamby 252
hamby252_results <- 
  readRDS("~/Research/CMPSpaper/preconsideration/hamby252_results.rds")

b252.full <- 
  readRDS("~/Research/CMPSpaper/preconsideration/bullets_hamby252_Sep9.rds")

# generate bullet_id
bulletid.tb <- b252.full %>% select(scan_id)
bulletid.tb <- bulletid.tb %>% mutate(
  barrel_id = sapply(strsplit(scan_id, "-"), "[[", 1),
  bullet = sapply(strsplit(scan_id, "-"), "[[", 2),
  land_id = sapply(strsplit(scan_id, "-"), "[[", 3),
  bullet_id = paste(barrel_id, bullet)
)

# remove x3p files
b252 <- b252.full %>% select(-x3p,-crosscut,-grooves,-ccdata)

# get bullet id
b252$bullet_id <- bulletid.tb %>% pull(bullet_id)

# get all comparisons
b.cb <- unique(b252$bullet_id) %>% utils::combn(m = 2)
p <- 1:595

# remove damaged data
tmp.tibble <- tibble(bullet1 = b.cb[1,], 
                     bullet2 = b.cb[2,])

tmp.tibble <- tmp.tibble %>% mutate(
  type = if_else(
    bullet1 %>% str_extract("Br\\d+") ==
      bullet2 %>% str_extract("Br\\d+"),
    "KM",
    "KNM"
  ),
  type = if_else(is.na(type), "Ukn", type)
)

# include tank-rashed data
t1.df <-
  read_rds("~/Research/CMPSpaper/preconsideration/t1.df.rds")
tr.id <-
  c(
    t1.df %>% filter(type == "TR") %>% pull(scan_id) %>% as.character(),
    "Br6-2-1",
    "Br9-2-4"
  )

######################################
# add ground-truth for hamby 252
studyinfo <-
  readxl::read_xlsx(
    "~/Research/CMPSpaper/preconsideration/StudyInfo.xlsx",
    sheet = 3,
    skip = 1
  )

keyss <- studyinfo$`Specimen ID`[-length(studyinfo$`Specimen ID`)]
keyss <- tibble(specimen = keyss)
keyss <- keyss %>% mutate(
  barrel = specimen %>% str_split(" "),
  barrel.id = purrr::map_chr(barrel, function(bb) {
    bb[1]
  }),
  bullet.id = purrr::map_chr(barrel, function(bb) {
    bb[3]
  })
) %>% select(-barrel)

# obtain the matching keys
keyss <- keyss %>% filter(!(bullet.id %>% str_detect("\\d+")))

tmp.ukn <-
  tibble(
    col1 = tmp.tibble$bullet1 %>% str_extract(" [A-Z]") %>% str_trim(),
    col2 = tmp.tibble$bullet2 %>% str_extract(" [A-Z]") %>% str_trim()
  )

# left join the ground truth
tmp.tibble$b1t <-
  tmp.ukn %>% left_join(keyss, by = c("col1" = "bullet.id")) %>%
  pull(barrel.id) %>% str_remove("Brl")
tmp.tibble$b2t <-
  tmp.ukn %>% left_join(keyss, by = c("col2" = "bullet.id")) %>%
  pull(barrel.id) %>% str_remove("Brl")

# find KM KNM
tmp.tibble <- tmp.tibble %>% mutate(
  b1t = if_else(
    is.na(b1t),
    bullet1 %>% str_extract("Br\\d+") %>% str_remove("Br"),
    b1t
  ),
  b2t = if_else(
    is.na(b2t),
    bullet2 %>% str_extract("Br\\d+") %>% str_remove("Br"),
    b2t
  ),
  b1t = as.numeric(b1t),
  b2t = as.numeric(b2t),
  type_truth = if_else(b1t == b2t, "KM", "KNM")
)

### set up parameters
N <- 4
CMPS_hamby252_results <- list()
CMPS_hamby252_results$span1 <- list(0.75, 0.25, 0.25, 0.15)
CMPS_hamby252_results$signame <-
  list("sigs75", "sigs25", "sigs25_1062", "sigs15")
CMPS_hamby252_results$npeaks.set <-
  list(c(5, 3, 1),
       c(5, 3, 1),
       c(10, 6, 2),
       c(5,3,1))
CMPS_hamby252_results$seg_length <- as.list(rep(50, 4))
CMPS_hamby252_results$Tx <- as.list(rep(25, 4))
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby252_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby252_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

for (i in 1:N) {
  b252[[CMPS_hamby252_results$signame[[i]]]] <- purrr::map2(
    .x = b252.full$ccdata,
    .y = b252.full$grooves,
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x,
        grooves = y,
        span1 = CMPS_hamby252_results$span1[[i]],
        span2 = 0.03
      )
    }
  )

  # process of removing outliers
  tt <-
    lapply(b252[[CMPS_hamby252_results$signame[[i]]]], function(x) {
      x$sig
    }) %>% unlist()
  qtt <- quantile(tt, na.rm = TRUE)
  multi.iqr <- 3
  outrange <-
    c(qtt[2] - multi.iqr * (qtt[4] - qtt[2]), qtt[4] + multi.iqr * (qtt[4] -
                                                                      qtt[2]))

  b252 <- b252 %>% mutate(sigs_main = purrr::map(
    .x = .data[[CMPS_hamby252_results$signame[[i]]]],
    .f = function(x) {
      tmp <- x$sig
      tmp[tmp < outrange[1] | tmp > outrange[2]] <- NA
      list(sig = tmp)
    }
  ))

  # start the parallel computing
  system.time({
    tmp.252.list <- mclapply(p, function(cb.idx) {

      tmp.lands <-
        c(
          b252 %>% filter(bullet_id %in% b.cb[, cb.idx][1]) %>% pull(scan_id),
          b252 %>% filter(bullet_id %in% b.cb[, cb.idx][2]) %>% pull(scan_id)
        )
      tmp.comp <-
        data.frame(expand.grid(land1 = tmp.lands[1:6], land2 = tmp.lands[7:12]),
                   stringsAsFactors = FALSE)

      tmp.comp$cmps <-
        purrr::map2(
          .x = tmp.comp$land1,
          .y = tmp.comp$land2,
          .f = function(xx, yy) {
            sig1 <- b252$sigs_main[b252$scan_id == xx][[1]]$sig
            sig2 <- b252$sigs_main[b252$scan_id == yy][[1]]$sig

            extract_feature_cmps(
              sig1,
              sig2,
              npeaks.set = CMPS_hamby252_results$npeaks.set[[i]],
              seg_length = CMPS_hamby252_results$seg_length[[i]],
              Tx = CMPS_hamby252_results$Tx[[i]],
              include = "nseg"
            )
          }
        )

      tmp.comp <- tmp.comp %>%
        mutate(
          cmps_score = sapply(tmp.comp$cmps, function(x)
            x$CMPS.score),
          cmps_nseg = sapply(tmp.comp$cmps, function(x)
            x$nseg)
        )

      cmps.table <-
        tmp.comp %>% select(land1, land2, cmps_score, cmps_nseg)

      tibble(
        bullet1 = b.cb[, cb.idx][1],
        bullet2 = b.cb[, cb.idx][2],
        cmps.table = list(cmps.table)
      )
    }, mc.cores = detectCores())
  })

  hamby252.cmps <- do.call(rbind, tmp.252.list)

  ##################################################
  # before removing tank-rashed data
  hamby252.cmps <- hamby252.cmps %>% mutate(
    cmps.max = cmps.table %>% purrr::map_dbl(
      .f = function(t) {
        max(t$cmps_score)
      }
    ),
    cmps.maxbar = cmps.table %>% purrr::map_dbl(
      .f = function(t) {
        landidx1 <-
          t$land1 %>% as.character() %>% str_sub(-1, -1) %>% as.numeric()
        landidx2 <-
          t$land2 %>% as.character() %>% str_sub(-1, -1) %>% as.numeric()
        compute_average_scores(landidx1, landidx2, t$cmps_score) %>% max()
      }
    )
  )

  hamby252.cmps$cmps.table.m <-
    lapply(hamby252.cmps$cmps.table, function(tt) {
      tt.idx <- tt %>% rowid_to_column %>%
        filter(land1 %in% tr.id | land2 %in% tr.id) %>% pull(rowid)
      tt[tt.idx, "cmps_score"] <- NA
      tt
    })

  # after removing tank-rashed data
  hamby252.cmps <- hamby252.cmps %>% mutate(
    cmps.max.m = cmps.table.m %>% purrr::map_dbl(
      .f = function(t) {
        max(t$cmps_score, na.rm = TRUE)
      }
    ),
    cmps.maxbar.m = cmps.table.m %>% purrr::map_dbl(
      .f = function(t) {
        landidx1 <-
          t$land1 %>% as.character() %>% str_sub(-1, -1) %>% as.numeric()
        landidx2 <-
          t$land2 %>% as.character() %>% str_sub(-1, -1) %>% as.numeric()
        compute_average_scores(
          landidx1, landidx2, t$cmps_score, addNA = TRUE
        ) %>% max()
      }
    )
  )

  # add type and type_truth
  hamby252.cmps$type <- tmp.tibble$type
  hamby252.cmps$type_truth <- tmp.tibble$type_truth

  ###################################
  # save result in a list
  CMPS_hamby252_results$cmps.table[[i]] <- hamby252.cmps

}

>>>>

The white horizontal line in \autoref{fig:process}(a) indicates which cross-section was identified by the algorithm to represent the LEA; \autoref{fig:process}(b) shows the corresponding cross-sectional view. 
Groove engraved areas are removed from the analysis as indicated by the vertical blue lines in \autoref{fig:process}(c) and \autoref{fig:process}(d)). A non-parametric LOESS smooth  [@loess] is fitted to capture the bullet curvature (\autoref{fig:process}(e)) and, finally, the **bullet signature** (\autoref{fig:process}(f)) is obtained as residuals of the cross-section and the fitted smooth. 
Note, that in @cmps bullet signatures are  referred to as bullet profiles. However, to avoid confusion, we distinguish the notion of bullet signatures from bullet profiles. Bullet profiles are shown in panels (b), (c), and (d) of \autoref{fig:process}, while \autoref{fig:process}(f) shows the corresponding bullet signature.
Identifying the groove engraved areas correctly is fundamental for a correct down-stream analysis of the signatures. To allow for a human-in-the middle inspection and intervention we have provided an interactive web-application, implemented as `R` Shiny App [@shiny],  `bulletinspectR` to identify and correct those errors. 
An example of the extraction process with corresponding code and parameter settings can be found on Github at \url{https://github.com/willju-wangqian/CMPSpaper/tree/main/reproducible}
Note that the process of extracting signatures might be different from the one used in @cmps because no code or parameter settings are made available publicly.

@@

### Conceptual idea of CMPS
Most algorithms for comparing striation marks are based on the digitized signatures and produce a similarity score [@song2005, @ChumbleyL_Scott2010VoTM, @aoas, @pmid30444940 ].
 The congruent matching profile segments (CMPS) algorithm, developed by @cmps for "objective comparison of striated tool marks", is one such algorithm. The algorithm's main idea  is to take a set of consecutive and non-overlapping basis segments from  the comparison  and  for each segment find the "best" registration position on the reference  (the other bullet signature) with respect to their cross-correlation values. From a comparison of these registration positions, a **congruent registration position** is identified, and the number of basis segments taking the congruent registration position is the CMPS score. \wj{XXX Note that researchers in} @cmps \wj{refers the origin of basis segments as the reference, but in this paper we refer it as the comparison.}
  High CMPS scores are achieved between more similar signatures and are therefore indicative of a same-source pair. Low scores between pairs of signatures are attributed to different source pairs.
\wj{However, a specific threshold of the CMPS score to distinguish between same-source and different-source comparisons is not mentioned in @cmps and depends on the underlying structure of the data and the choice of parameters. Further research is needed to understand how to determine the threshold. XXX}
  <!-- \hh{XXX What threshold did @cmps identify as the boundary between different source and same source? And what was the error rate in their example?} -->
  <!-- \wj{they didn't talk about a specific threshold or error rates} \hh{XXX Is there a way to infer the error based on the example? ... 'while the authors did not mention a specific threshold of the CMPS to distinguish between same-source and different-source comparison, a CMPS value of Y results in an error rate of Z in the example.'} -->
 The CMPS algorithm can assist firearm examiners with drawing a conclusion about the source of a comparison pair. Unfortunately, @cmps did not release any code or specific parameter settings for their implementation of the CMPS algorithm. 
Thus, in this paper we present an open-source implementation of the CMPS algorithm in the R package CMPS available from Github \url{https://github.com/willju-wangqian/CMPS}
With our implementation, one can compute the CMPS score of a comparison using the following code:

@@

1. **cut the comparison signature into consecutive, non-overlapping and equal-length basis segments:**
The command `get_segs(x, len=50)` implements this step:  it takes  bullet signature `x` in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`, which are referred to as "basis segments".

2. **identify candidate positions:** For each basis segment a set of candidate registration  positions on the comparison signature is identified based on the segment's similarity to the reference signature. In a first step, the cross-correlation function of the segment to the reference is calculated, then a number of positions with high correlation values are identified as candidate positions. In case multiple segment lengths are considered, the length of each basis segment is doubled and these two steps are repeated. Only when candidate positions coincide (or are similar enough), they are considered further. \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.
    + **Calculate the cross-correlation curve:**  Calculate the cross-correlation curve between a basis segment `x` and the reference signature `y` using the function `get_ccf4(x, y, ...)` as shown in \autoref{fig:segplot2}. The position indicates the lag by which a basis segment is moved with respect to its original placement. A position is considered "good" if it results in a peak in the cross-correlation between the basis segment and the reference.

@@

        -   `get_ccr_peaks(comp, segments, seg_scale, nseg = 1, npeaks = 5)` computes the cross-correlation curve between a basis or scaled segment and the reference signature and finds peaks in the cross-correlation curve. The number of peaks detected is equal to `npeaks`, which is an integer. `segments`, `seg_scale`, and `nseg` determines the segment in the cross-correlation computation, and `comp` gives the reference signature. If the multi-segment lengths strategy is used, then `get_ccr_peaks(...)` is called in a `lapply()` for each level of the basis segment. The resulting list is called `ccr.list`.

@@

3. **determine the congruent registration position:** A candidate position "receives" `votes` from basis segments that identify it or a close position within a tolerance zone of `Tx` as a candidate position in step 2. Votes for all candidate positions are tallied, and the position with the highest number of votes gets chosen as the *congruent registration position*,  indicating that most of the basis segments find their highly similar counterpart in the reference signature in terms of correlation at this registration position. In case of ties the middle position is taken as the congruent registration position. Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score of the comparison. `get_CMPS(input.ccp, Tx = 25)` is the function that tallies the votes and determines the congruent registration position and congruent matching profile segments (cmps).

@@

* `x` and `y` are  two  signatures:  `x` serves as the comparison signature (which will be divided into basis segments) and `y` is the reference signature;

* `include` determines the \hh{format of the function result}. Besides the `CMPS.score`, other aspects of the comparison help in understanding how the `CMPS.score` is computed. By default `include` is set to `NULL` and  only  the `CMPS.score` is returned; further results are included when `include` is (an abbreviation of) one of or a vector of the following strings: `"nseg", "congruent.pos", "congruent.seg.idx", "segments", "parameters"`, and `"full_result"`. If `include` is specified as `"full_result"` (or its abbreviation), the ouput includes everything listed below.
  - `nseg`: the number of basis segments  from the comparison signature; this is also the highest possible CMPS score of the comparison;
  - `congruent.pos`: the congruent registration position;
  - `congruent.seg.idx`: the indices of all congruent matching profile segments;
  - `ccp.list`: a list showing identified candidate positions of all basis segments; 
  - `pos.df`: a data frame containing all candidate positions and their respective number of votes;
  - `segments`: a list containing all basis segments;
  - `parameters`: a list containing all input arguments of `extract_feature_cmps`;

@@

The `CMPS` package contains a simple example to illustrate the basic usage of the package. The data in this example are twelve bullet signatures obtained from two bullets in Hamby set 252 [@hamby].  The procedure for generating signatures from high-resolution 3D topographic scans of bullet lands used here follows the methodology described in @aoas (as discussed above). 
The two bullets under consideration are known to have been fired from the same gun barrel, so for the 36 pairwise land-by-land comparisons, six comparisons are from same-source pairs (known matches) while thirty are from different-source pairs (known non-matches). To access the example data, we use

@@

The signatures of Land 4 of Bullet 1 and Land 5 of Bullet 2 are stored in objects `sigs2` and `sigs1`, respectively. This comparison consists of a pair of signatures that are known to be a match -- a KM (known match) comparison. We compute the CMPS score using two versions of the CMPS algorithm:

@@

We also implemented graphing tools for visualizing results of the CMPS algorithm. The goal is to provide users with tools to inspect each of the basis segments and to help them have a better understanding of how the algorithm works. \autoref{fig:sigplots} shows the plots generated by the first graphing function, `cmps_signature_plot()`, and continues with the example above. `cmps_signature_plot()` takes the output of `extract_feature_cmps(..., include = "full_result")` and returns a list of 5 elements. It creates an overall impression of how the comparison signature aligns with the reference signature at the congruent registration position.

@@

*   The second plot is called `signature_shift_plot`, shown in \autoref{fig:sigplot2}. This visual presents both the comparison signature and the reference signature. The comparison signature is aligned with the reference signature based on the congruent registration position. Congruent matching profile segments are highlighted by  solid red lines.

\begin{figure}[hbt]
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot}The black line shows the comparison signature; each red line segment shows one congruent matching profile segment.}
```{r sigplot, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
sig.plot <- cmps_signature_plot(
  cmps_with_multi_scale
)
sig.plot$segment_shift_plot
```
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot2}The black line shows the reference signature; the red line shows the comparison signature. Solid part shows the congruent matching profile segments, and the dashed part shows segments that do not agree with the congruent registration position.}
```{r sigplot2, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
sig.plot$signature_shift_plot
```
\end{subfigure}
\caption{\label{fig:sigplots} Two visual outputs of cmps\_signature\_plot()}
\end{figure}

@@

*   \autoref{fig:segplot} is the `segment_plot` for basis segment 6 at level one (in its original length). We used  `npeaks.set = c(5, 3, 1)` in `extract_feature_cmps()` when calculating the CMPS score. Therefore the top five peaks are identified in the cross-correlation curve at level one. Segment 6 is plotted at the positions where these 5 peaks are identified with dashed lines in the `segment_plot`. The solid thick black line shows the segment at its original position (which in this example is very close to the actual registration position). 
*   \autoref{fig:segplot2} is the `scale_ccf_plot` of basis segment 6 at level one. It shows the cross-correlation curve computed by the reference signature and the level-one basis segment 6. The five highest peaks are marked by dots on the curve.

\begin{figure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:segplot}segment\_plot for segment 6 at level one. The original position of segment 6 is indicated by the solid black line. Positions where the segment achieves the 5 highest cross-correlations are indicated by the dashed line segments.}
```{r segplot, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
seg.plot <- cmps_segment_plot(
  cmps_with_multi_scale, 
  seg.idx = 6
)
seg.plot[[1]]$segment_plot
```
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:segplot2}scale\_ccf\_plot shows the cross-correlation curve between the reference signature and segment 6 at level one. The five highest peaks are marked by dots. The vertical red dashed line indicates the congruent registration position; the green dashed line shows a peak position in the highest segment level; the blue dashed lines show the tolerance zone around the green dashed line. We can see that none of the five highest peaks at level one falls within the tolerance zone, indicating that there is no consistent correlation peak or a candidate position identified by basis segment 6 under the multi-segment lengths strategy. Thus, the basis segment 6 doesn't vote for the congruent registration position and is not a cmps.}
```{r segplot2, opts.label="codefig", echo=TRUE, out.width='\\textwidth', warning = FALSE}
seg.plot[[1]]$scale_ccf_plot
```
\end{subfigure}
\caption{\label{fig:segplots} Two plots used to investigate basis segment 6 at level one}
\end{figure}

@@

ggarrange(
  plotlist = 
    unlist(seg.plot, 
           recursive = FALSE),
  ncol = 2, 
  nrow = 3)
```
\caption{\label{fig:seg_all} Put segment\_plot and scale\_ccf\_plot of all three levels together. We are identifying five highest peaks at level one, three peaks at level two, and one peak at level three since npeaks.set = c(5, 3, 1). The highest peak position at level three is marked by the green dashed line across all segment levels. However, the highest peak on level three does not coincide with any of the top five highest peaks at level one. This indicates that there is no consistent correlation peak or a candidate position for basis segment 6 under the multi-segment lengths strategy.}
\end{figure}

@@

In general, we assume each bullet has $n$ land engravings (in our case $n=6$). Let $c_{ij}$ denote the CMPS score of a comparison between bullet 1 land $i$ and bullet 2 land $j$, for $i,j = 1, \dots, n$. Let $\mathcal{P}_k$ denote bullet land pairs in phase $k$ for $k = 0, \dots, n-1$, and $\mathcal{P}_k = \{ (i,j): i = 1, \dots, n ; j = (i + k) - n \cdot \mathbbm{1}_{i + k > n} \}$, where $\mathbbm{1}_A$ denotes an indicator function. For example, $\mathcal{P}_1 = \{ (1,2), (2,3), (3,4), (4,5), (5,6), (6,1) \}$ when $n = 6$.

@@

The codes generating those results can be found on Github at \url{https://github.com/willju-wangqian/CMPSpaper/tree/main/reproducible}

@@
